Supplementary material: The increasing cost of happiness

### 2021-06-10 src/supplementary-SSM1.Rmd
Closes issue #21
Update to discussion of Table S3. Closes issue #17

@@

The posterior estimates of the piecewise model indicated that the relationship between happiness and income changed over time between 2001 and 2019. One implication of such a change is the disparity in happiness between income groups has increased. This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and happiness are strongly related. To directly examine the implications of our model for such inequities in happiness, we used the model to estimate or _predict_ the happiness of two different income levels: one income level which was above the change point in 2001 but fell below it by 2019 (\$50K/yr); and another level which always remained above the change points over the same period (\$75K/yr). Figure S8 below presents the happiness levels (in SD units/year) for each income level as well as the difference in happiness between them (∆).

>>>>

The table of model parameters (untransformed) for each year show that the $\beta_1$ slopes for piecewise-linear, log-linear and linear model for affective wellbeing and income do not show a sustained trend ove the 19 year period. The $\beta_1$ coefficients increase in the initial 5-10 years and then decline. Despite this common pattern, the linear slope appears to show the most sustained increase while the log-linear slope shows a relatively less sustained increase. Overall, the pattern of changes among the models indicates the function between income and happiness has in fact become slightly steeper over the period.

@@

The posterior estimates of the piecewise model indicated that the relationship between happiness and income changed over time between 2001 and 2019. One implication of such a change is the disparity in happiness between income groups has increased. This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and happiness are strongly related. To directly examine the implications of our model for such inequities in happiness, we used the model to estimate or _predict_ the happiness of two different income levels: one income level which was above the change point in 2001 but fell below it by 2019 (\$50K/yr); and another level which always remained above the change points over the same period (\$75K/yr). The figure below presents the happiness levels (in SD units/year) for each income level as well as the difference in happiness between them (∆).
